{\tt APFELgrid}: a high performance tool for parton density determinations

### 2016-05-05 paper.tex
extra references

Measurements at the Large Hadron Collider (LHC) have a unique capacity to shed light upon the internal dynamics 
of the proton and provide new constraints upon proton PDFs. However the inclusion of large collider datasets 
into PDF fits can provide a significant challenge due to the large computational footprint of performing accurate
theoretical predictions over the many iterations required in a fitting procedure. In order to make the fullest use of current 
and future LHC results, efficient strategies for the computation of these observables must therefore be employed.

@@

The {\tt FastKernel} method was developed to
address this problem in the context of the NNPDF global analyses~\cite{Ball:2010de,Ball:2014uwa}. This method differs from the standard procedure
\`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
amount of information that is precomputed prior to fitting so as to
minimise the amount of operations required during the
fit. More specifically, the {\tt FastKernel} method relies on the
combination of precomputed hard cross sections with DGLAP evolution
kernels into a single look-up table, here called
a {\tt FastKernel} ({\tt FK}) table. In this way the prediction
for a given hadronic observable can be obtained by performing a simple matrix
product between the respective {\tt FK} table and PDFs evaluated directly at the fitting scale.

@@

We shall now examine the performance differences between
the two expressions for fast interpolated cross section prediction
Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
  FK}).
%
In order to provide a comprehensive benchmark, we consider here a wide range of
processes including LHC and Tevatron electroweak vector boson
production
measurements~\cite{Aaij:2012mda,Aaij:2012vn,Chatrchyan:2013mza,Chatrchyan:2013uja,Chatrchyan:2012xt,Aad:2013iua,Aad:2011fp,Aad:2011dm,Aaltonen:2010zza},
$t\bar{t}$ total
cross sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
double-differential Drell-Yan
cross sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
are performed over a wide range of kinematics, for a total of 52
source {\tt APPLgrid} files corresponding to the majority of available
LHC and Tevatron datasets of applicability to PDF determination.
While the source {\tt APPLgrid} files have varying grid densities in
$x$ and $Q^2$, for the purposes of comparison the corresponding {\tt
  FK} tables are produced consistently with 30 points in $x$, and at
an initial scale below the charm threshold, therefore with seven
active partonic species.

>>>>

Measurements at colliders such as the Tevatron and Large Hadron Collider (LHC) have a unique capacity to shed light upon the internal dynamics 
of the proton and provide constraints upon proton PDFs~\cite{Rojo:2015acz}. However the inclusion of large hadron collider datasets 
into PDF fits can provide a significant challenge due to the large computational footprint of performing accurate
theoretical predictions over the many iterations required in a fitting procedure. In order to make the fullest use of current 
and future LHC results, efficient strategies for the computation of these observables must therefore be employed.

@@

The {\tt FastKernel} method was developed to
address this problem in the context of the NNPDF global analyses~\cite{Ball:2014uwa}. This method differs from the standard procedure
\`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
amount of information that is precomputed prior to fitting so as to
minimise the amount of operations required during the
fit. More specifically, the {\tt FastKernel} method relies on the
combination of precomputed hard cross sections with DGLAP evolution
kernels into a single look-up table, here called
a {\tt FastKernel} ({\tt FK}) table. In this way the prediction
for a given hadronic observable can be obtained by performing a simple matrix
product between the respective {\tt FK} table and PDFs evaluated directly at the fitting scale.

@@

We shall now examine the performance differences between
the two expressions for fast interpolated cross section prediction
Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
  FK}).
%
In order to provide a comprehensive benchmark, we consider here a wide range of
processes including LHC and Tevatron electroweak vector boson
production
measurements~\cite{Aaij:2012mda,Aaij:2012vn,Chatrchyan:2013mza,Chatrchyan:2013uja,Chatrchyan:2012xt,Aad:2013iua,Aad:2011fp,Aad:2011dm,Aaltonen:2010zza},
$t\bar{t}$ total
cross sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
double-differential Drell-Yan
cross sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
are performed over a wide range of kinematics, for a total of 52
source {\tt APPLgrid} files corresponding to the majority of available
LHC and Tevatron datasets of applicability to PDF determination.
While the source {\tt APPLgrid} files have varying grid densities in
$x$ and $Q^2$, for the purposes of comparison the corresponding {\tt
  FK} tables are produced consistently with 30 points in $x$, and at
an initial scale below the charm threshold, therefore with seven
active partonic species. These settings are chosen so as to provide a realistic
comparison, in a production environment the density and distribution of the
$x$-grid may be adjusted to match interpolation requirement.
